---
title: "Revision of hospitalization and discontinuation paper"
output: html_notebook
editor_options: 
  chunk_output_type: inline
---

```{r setup, include = F}
suppressPackageStartupMessages(library(tidyverse))
library(data.table)
library(fs)
library(glue)
condition_names <- c('stroke_primary' = 'Primary stroke',
                     'stroke_compl' = 'Complicated stroke',
                     'LuCa' = 'Lung cancer',
                     'MetsCa' = 'Metastatic cancer',
                     'dement' = 'Dementia',
                     'thrive' = 'Failure to thrive')
```


## Source files

- hospital_comorb.R
- hospitalization.R
- cause_esrd.R
- hospitalization_TabFig.R

## Required revisions

1. Specification of comorbidity index and inclusion of cause of ESRD
1. Rates of transplant in each of the cohorts
1. Test the inclusion of cause of ESRD in all the analyses, starting with stroke, and
   see if there are any differences
1. Potential pitfalls of the simulation model, and model fit
1. Better specification of the simulation model
1. Sensitivity analysis ignoring transplantation

# Analyses addressing each revision point

## Specification of comorbidity index and inclusion of cause of ESRD

The comorbidity score is based on Liu, et al (2010) and is computed as

```
comorb_indx = ASHD + 3 * CHF + 
  2 * (CVATIA + PVD + Other.cardiac + COPD + GI.Bleeding + Liver + 
       Dysrhhythmia + Cancer) +
  Diabetes
```

## Rates of transplant

```{r}
drop_dir <- path(ProjTemplate::find_dropbox(),'NIAMS','Ward','USRDS2015')
hosp_data <- readRDS(path(drop_dir, 'data','final_hosp_data.rds'))
map(hosp_data, 
    ~summarize(., Total = sum(cens_type==2), Percentage = 100*mean(cens_type==2))) %>% 
  bind_rows(.id = 'Condition') %>% 
  mutate(Condition = condition_names[Condition],
         Percentage = round(Percentage, 1)) %>% 
  knitr::kable()
```

## Test the inclusion of cause of ESRD in all the analyses, starting with stroke, and see if there are any differences



## Potential pitfalls of the simulation model, and model fit


## Better specification of the simulation model


## Sensitivity analysis ignoring transplantation
