---
title: "Simulation Study"
author: "Abhijit Dasgupta"
date: "7/12/2019"
output: 
  word_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Objective

We have postulated that the observed racial difference in survival in ESRD patients 
can be attributed to different propensities to discontinue dialysis. In this study, 
we will model the dialysis discontinuation process in whites as a Weibull regression model,
and use it to provided predicted discontinuation times for other racial groups; this will 
represent discontinuation times in those racial groups if they behaved the same as whites. We will
then use these discontinuation times to compute survival times for each subject, i.e. the minimum of 
the time of death, the time of transplant, the predicted time of discontinuation under the Weibull model for whites, and loss-to-followup. For mortality, we will consider death or discontinuation as 
the event of interest (since median survival time after discontinuation is 7 days), with transplant and loss to followup being censoring events. 

## Implementation

In this simulation
study, for each patient subgroup based on disease condition (stroke, lung
cancer, metastatic cancer, dementia, failure to thrive) we first model the time
to discontinuation in whites as a Weibull regression model, with death,
loss-to-followup and kidney transplant being censoring events and dialysis
discontinuation being the event of interest. We call this model the whites
model. The regressors (independent variables) are age at the time the disease
condition was diagnosed (categorized by decade of age), gender, time on
dialysis, geographic region (Northeast, Midwest, South and West), socio-economic
status score, comorbidity index and cause of ESRD (Diabetes, Glumerulonephritis,
Hypertension and Other). The models were fit using the `survreg` function from
the `survival` package (version 2.44.1.1) in R (version 3.6). Note here that
even for the same combination of covariate values, different people will
experience discontinuation at different times; the underlying assumption we make
is that these times arise from some Weibull distribution (which is a flexible
family of probability distributions commonly used to model survival times). The
whites model gives us, for every combination of covariate values, the estimated
parameters for the Weibull distribution from which the corresponding
discontinuation time arose. This means that we can simulate discontinuation times
for any individual based on the whites model by simulating data from the Weibull
distribution corresponding to the particular combination of covariate values that 
individual has according to the model.

Let us now consider the data from blacks. Each individual has observed times for
death, discontinuation, transplant and loss-to-followup (LTF), along with values of 
each of the measured covariates. For each individual, we replace the observed 
discontinuation time with a random discontinuation time simulated from the Weibull
distribution determined from the whites model; this is a simulation of when the individual 
would have discontinued if he or she behaved like a white person with similar characteristics. We then re-evaluate the individual's status. 
If the simulated discontinuation time is before the observed death, transplant and LTF times, then
we record the individual's status as "discontinued"; otherwise we record the status according as whichever
of the observed death, transplant and LTF times was earliest. This can change 
individuals' status to "discontinued" from their observed status or change them
from an observed "discontinued" status to something else, or leave their status unchanged
compared to what was observed. Correspondingly, the time they are in the study can also change, and
will be the earliest of observed time of death, simulated time of discontinuation, 
observed time of transplant and observed LTF time. We repeat this process 1000 times, to 
allow the variability in the simulated discontinuation times to be expressed

Let's try to understand this with an example. Let's consider a 70 year old black man 
from the South who never received a transplant and died 5 years after starting dialysis, with 
some combination of other covariate values. We input his covariate values into the 
whites model, and get an estimate of the Weibull distribution parameters, from which we
simulate a discontinuation time of 4.5 years. Since this is earlier than his 
observed death time of 5 years, within the simulation study his status changes to 
"discontinued" and his time in the study changes to 4.5 years. The next time, we simulate
a discontinuation time of 5.2 years, in which case this man's status remains "dead" and 
his time in the study remains 5 years. We do this 1000 times, each time simulating a new
discontinuation time based on the whites model and re-assessing his status and time in study. 

For each of the 1000 rounds in the simulation study, we do this process for each individual
in the black group. We also do the exact same process with each individual in the Asian, Hispanic, 
and Native American groups. We then pool all the race groups together and fit a Cox regression
with death or discontinuation as the event of interest and race as the only covariate. This gives us estimated hazard ratios for each race compared to whites, under the simulation condition that
all individuals behave like similar whites in terms of discontinuation. For a 1000 rounds
of simulation we get 1000 hazard ratios comparing each of the other races to whites. This
gives us a distribution of hazard ratios simulated under our condition that all races behave
just like similar whites with respect to discontinuation. We then compare this distribution
to the __observed__ hazard ratios comparing each race to whites. If the pattern of 
discontinuation was not a factor in mortality, we would expect the observed hazard 
ratios to fall within the distribution of the simulated hazard ratios. However, Figure 2 shows
that not to be the case for blacks, Hispanics and Asians. 

## Diagnostics

This simulation study depends on whether we have adequately been able to model the
discontinuation process in whites using our whites model. One way to see the 
adequacy of model fit in survival models is with Cox-Snell residuals, which should
follow an Exponential distribution with parameter 1 if the model is a good fit. In 
the presence of censoring, these residual values are also censored, so we can 
estimate their distribution using a Kaplan-Meier estimate, and see how well it 
corresponds to the survival function (1 - the cumulative distribution function) of
the unit Exponential distribution. The figure below shows the distribution of the 
Cox-Snell residuals against the exponential distribution (red line), and shows
adequate fit for each of the disease groups.

```{r,echo=F, message = F, warning=F}
dropdir <- file.path(ProjTemplate::find_dropbox(), 'NIAMS','Ward','USRDS2015','data')
abhiR::reload()
load(file.path(dropdir,'revision_JASN', 'modeling_data.rda'))
weib_models <- list()
for (cnd in names(modeling_data2)){
  D = modeling_data2[[cnd]]
  weib_models[[cnd]] <- survreg(Surv(time_from_event+0.1, cens_type==3)~ # Added 0.1 since weibull is > 0
                                  agegrp_at_event + SEX  + time_on_dialysis +
                                  REGION+
                                  zscore + comorb_indx + ESRD_Cause,
                                data = D$White,
                                dist = 'weibull')
}
condition_code <- c('stroke_primary' = 'Primary Stroke',
                    'stroke_compl' = 'Complicated Stroke',
                    'LuCa' = 'Lung Cancer',
                    'MetsCa' = 'Metastatic Cancer',
                    'dement' = 'Dementia',
                    'thrive' = 'Failure to thrive'
                    )
par(mfrow = c(2,2))
for(n in c('stroke_primary','LuCa','dement','thrive')){
  cs <- cox_snell(weib_models[[n]], modeling_data2[[n]]$White)
  title(main = condition_code[n])
}
```
